---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  comment = ""
)
devtools::load_all()
options(width = 100)
```

# FamModel

<!-- badges: start -->
<!-- badges: end -->

This package implements regression models for family-based genetic studies,
including mixed models. While such models can be fit with other existing
software, this package tries to provide a unified solution that leverages
the flexibility of the R environment. It also leverages modern automatic
differentiation techniques to fit likelihood-based models quickly.

## Installation

You can install the latest released version of FamModel from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kinnamon-lab/FamModel", ref = "[tag]")
```

where `[tag]` is the tag of the most recent release.

## Example

The following code fits a univariate version of the linear mixed model presented
in [Cowan et al. (2018)](https://doi.org/10.1161/CIRCGEN.117.002038) with an
equivalent parameterization using the `lmna_nonseg` example data provided with
the package. Note how the package permits flexible use of `formula` constructs
and provides an appropriate likelihood ratio test for a null narrow-sense
heritability on the boundary of the parameter space.

```{r example}
lmna_data <- copy(lmna_nonseg)[,
  `:=`(
    female = as.integer(sex == 2),
    age_echo_std = (age_echo_yrs - mean(age_echo_yrs, na.rm = TRUE)) /
      sd(age_echo_yrs, na.rm = TRUE)
  )
]
lmna_fd <- FamData$new(
  lmna_data,
  family_id = "family_ID",
  indiv_id = "individual_ID",
  proband = "proband",
  sex = "sex",
  maternal_id = "maternal_ID",
  paternal_id = "paternal_ID",
  mzgrp = "mzpair",
  dzgrp = "dzpair"
)
lmna_lvef_model <- lmna_fd$lmm(
  lvef ~ age_echo_std + female + I(n_lmna_vars > 0) + I(n_oth_vars > 0)
)
lmna_lvef_model$print()
```

## Acknowledgements

Development of this software was supported by the National Heart, Lung, and
Blood Institute and National Human Genome Research Institute of the National
Institutes of Health under award numbers R01HL128857 and R01HL149423. The
content is solely the responsibility of the author and does not necessarily
represent the official views of the National Institutes of Health.
